---
title: "Statistical Inference Course Project Report Part 1"
author: "Weihua Wang"
date: "Tuesday, January 14, 2015"
output: html_document
---

### Brief Introduction

This is the project report for the Online Coursera course -- Statistical Inference. Basic tasks of this project consists two parts:

1. A simulation exercise.
2. Basic inferential data analysis.

The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. 

For all of the thousand simulations, lambda should be set as 0.2. The distribution of averages of 40 exponentials should be investigated. 

### Simulation and results

One thousand simulated averages of 40 exponentials can be realized via creation of an empty vector to store all of our sample means.

```{r}
lambda = 0.2
n = 40
myexpd <- rep(NA, 1000)
```

Thus, the distribution of the mean of 40 exponentials is simulated by,

```{r}
set.seed(512)
for (i in 1:1000){
    tmp <- rexp(n, lambda)
    myexpd[i] <- mean(tmp) 
}
```

The mean and variance can be obtained by

```{r}
myexpd_m <- mean(myexpd)
myexpd_m
myexpd_v <- var(myexpd)
myexpd_v
```

Compared to the theoretical mean and variance

```{r}
mean_t <- 1 / lambda
mean_t
var_t <- (1/lambda)^2/n
var_t
```

As can be seen that the comparision between both the calculated mean/variance and the theoretical mean/variance featured with almost negligible variation.

To see if the distribution resembles a normal distribution.
```{r histogram}
hist(myexpd, probability=T, main="", xlab=""，col="blue", ylim=c(0,0.5))
lines(density(myexpd))
curve(dnorm(x,myexpd_m, sqrt(myexpd_v)), 2, 8, col="red", add=T)
```

Convert to data frame for ggplot.
```{r}
df <- data.frame(myexpd)
library(ggplot2)
```


```{r, echo=TRUE, fig.width=4, fig.height=3}
g <- ggplot(data = df, aes(x = myexpd)) 
ggp <- g + geom_histogram(aes(y=..density..), fill = I('blue'), binwidth = 0.20, color = I('black')) + stat_function(fun = dnorm, arg = list(mean = mean_t, sd = sqrt(var_t)), color = "red")
print(ggp)
```

Obviously, the distribution is approximately normal.

Last, the coverage of the confidence interval for 1/lambda is claculated through,

```{r}
lowercl <- myexpd - qnorm(0.975) * (1/lambda)/sqrt(n)
uppercl <- myexpd + qnorm(0.975) * (1/lambda)/sqrt(n)
expci <- mean(lowercl < (1/lambda) & uppercl > (1/lambda))
expci
```
The confidence interval is thus expci of 95.3%.
